---
title: "Linear Algebra"
output: html_notebook
---

## Section 1 - Scalars Vectors and Matrices

- Scalar - a constant
- Vector - a single row or column of numbers
- Matrix - a rectangular collection of numbers

- Dimensions - refer to the number of rows and columns
  - an rxc matrix - r rows and c columns
  - The m~ij~ element is at row i and column j
  
- Vectors are a subset of matrices
  - Either dimensions 1xc (a row vector) or rx1 (a column vector)
  - And scalars are 1x1
  
- Vectors are used to represent a location in space. Each entry represents a dimension. This is easy to see in 1, 2 and 3 dimensions
- Vectors are arrows, whenever we get beyond a single dimension

## Section 2 - Application Scalars Vectors and Metrics

1. Defining a vector [c(), :, seq(), rep()]
- R is fully vectorized - default process is to think in terms of vectors (i.e. a collection of elements)

```{r}
# Using the c() function
v1 <- c(2, 3, 4, 5)
str(v1)
# Using the ':' operator
v2 <- 2:5 # From 2 to 5, default counter 1

# USing the seq() function; a long-form version of the ':' operator
v3 <- seq(from = 2, to = 5, by = 1)

# OR...
v3 <- seq(2, 5, 1)

# Using the rep() function
v4 <- rep(37, 5)  # Repeats 37 5x times
```

2. Defining a matrix [matrix(), rbind(), cbind()]
- A matrix is just a collection of vectors
```{r}
# Using the matrix() function
m1 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = TRUE)
# byrow = TRUE - takes arguments and populates rowwise
m1
# Default is FALSE
# Useful to do it rowwise so can think rows first
str(m1)

# Using rbind()
m2 <- rbind(c(1, 2, 3), c(4, 5, 6))
m2
# Using cbind()
m3 <- cbind(c(1, 4), c(2, 5), c(3, 6))
m3
```

3. Plots in 2D and 3D
- Vector in 2-space
```{r}
# Vector in 2-space
u1 <- c(2, 4)
plot(u1[1], u1[2], type = 'p', col = 'red',
     xlim = c(0, 5), ylim = c(0,5),
     xlab = "", ylab = "")  
arrows(x0 = 0, y0 = 0,
       x1 = u1[1], y1 = u1[2],
       lwd = 2)  
# *0 - Tells where arrows should start
# Use reference to tell plot elements
```

- Vector in 3-space
```{r}
library(plot3D)
u2 <- c(2, 3, 4)
points3D(x=u2[1], y=u2[2], z=u2[3], col = 'red',
         xlim = c(0,5), ylim = c(0,5), zlim = c(0,5),
         phi = 0)  # Changes view
arrows3D(x0 = 0, y0 = 0, z0 = 0, x1 = u2[1], y1 = u2[2], z1 = u2[3],
         lwd = 2,
         add = TRUE)  # Adds to an existing plot - needed for 3D library
```

## Section 2 - LAB1 - Intro Scalars Vectors and Matrices
1. Create a 3x4 matrix, A, fill row-wise with the number 1-12. What is a~24~?
```{r}
A <- matrix(seq(1,12,1), nrow = 3, ncol = 4, byrow = TRUE)
A
A[2,4]
# a_24 = 8
```

2. Create two 1x2 vectors, v1 and v2, with numeric entities of your choice. Plot v1 and v2 on the same coordinate plane
```{r}
v1 <- c(5, 6)
v2 <- c(1, 3)
# Plot begins a session
plot(v1[1], v1[2], type = 'p', col = 'red',
     xlim = c(0,8), ylim = c(0,8),
     xlab = "", ylab = "")
arrows(x0 = 0, y0 = 0,
       x1 = v1[1], y1 = v1[2],
       lwd = 2, col = 'red')
# Points adds additional points
points(v2[1], v2[2], col = 'blue',
     xlim = c(0,8), ylim = c(0,8),
     xlab = "", ylab = "")

arrows(x0 = 0, y0 = 0,
       x1 = v2[1], y1 = v2[2],
       lwd = 2, col = 'blue')
```

3. Now add 3 to the second entry in v1 and re-plot. Describe the position of the new vector in 2-space relative to v1
```{r}
v2_b <- c(v1[1], v1[2] +3)
v2_b
plot(v1[1], v1[2], type = 'p', col = 'red',
     xlim = c(0,10), ylim = c(0,10),
     xlab = "", ylab = "")
plot(v2_b[1], v2_b[2], type = 'p', col = 'blue',
     xlim = c(0,10), ylim = c(0,10),
     xlab = "", ylab = "")
arrows(x0 = 0, y0 = 0,
       x1 = v1[1], y1 = v1[2],
       lwd = 2, col = 'red')
arrows(x0 = 0, y0 = 0,
       x1 = v2_b[1], y1 = v2_b[2],
       lwd = 2, col = 'blue')
```

4. Create a 1x3 vector, v3, with numeric entries of your choice. Plot v3 in 3-space
```{r}
v3 <- c(4, 5, 6)
points3D(x=v3[1], y = v3[2], z = v3[3], col = "red",
         xlim = c(0,8), ylim = c(0,8), zlim = c(0,8))
arrows3D(x0 = 0, y0 = 0, z0 = 0,
         x1 = v3[1], y1 = v3[2], z1 = v3[3],
         lwd = 2, add = TRUE)
```

5. Fill in the blanks: A *vector* is a *matrix* with only one *row* or *column*

## Section 3 - Vector operations
- **Scalar multiplication** - a scalar times a vector
  - Multiply each element of the vector by the scalar
  - 3 x [-1, 2, 5] = [-3. 6, 15]
  - The original vector is being **dilated** (stretched/compressed) by a factor equal to the scalar
  - If scalar > 1, then stretched
  - If scalar < 1, then compressed
  - If <0, then the direction will be flipped
  
- **Vector addition** - adding two vectors
  - Both vectors must have exactly the same dimensions
  - Then we add element wise
  - e.g. [2, 3] + [1, 4] = [3, 7]
  - The second vector is being drawn from the end of the first vector to create a new vector
  
- **Vector subtraction** - simply add the negative vector
  - Multiply each element of the second vector by the scalar -1, then subtract
  - v1 - v2 = v1 + -1*v2
  - Net effect is the same as subtracting the second vector from the first element-wise
  - r.h. [2, 3] - [1, 4] = [1, -1]
  - Illustrates a more general rule:
    - Multiplication and addition can be applied successively
    - v1 + s*v2
    - The order of operations is the same as with real number - multiply first, then add/subtract
    - Multiple each element of the second vector by the scalar then subtract
    - e.g. [2, 3] + 3 * [1, 4] = [5, 15]
    
- Linear algebra - in general, don't assume that rules in the real-number world still apply. The above is an exception

## Section 4 - Application vector operations

1. Plotting vectors
#### Plotting a single vector in 2-space

```{r}
vecplot_2d <- function(vector, xax, yax, color) {
  vec <- vector
  plot(vec[1], vec[2], type = 'p', col = color,
  xlim = xax, ylim = yax,
  xlab = "", ylab = "")
  arrows(x0 = 0, y0 = 0, x1 = vec[1], y1 = vec[2], lwd = 2, col = color)
  text(x = vec[1], y = vec[2], 
  labels = paste('(', vec[1], ',', vec[2], ')', sep=""),
  pos = 3)  # Pos 3 is to the right
  }
```

#### Apply the function
```{r}
vecplot_2d(vector = c(1,2), xax = c(0,4), yax = c(0,5), color="red")
```

#### Plotting a vector in 3-space
```{r}
library(plot3D)
vecplot_3d <- function(vector, xax, yax, zax, color) {
  vec <- vector
  points3D(x = vec[1], y = vec[2], z = vec[3],
           xlim = xax, ylim = yax, zlim = zax,
    xlab = "x", ylab = "y", zlab = "z")
text3D(x = vec[1] + 0.5, y = vec[2], z = vec[3],
       labels = paste("(",vec[1], ",", vec[2], ",", vec[3],")", sep = ""),
       add = TRUE)
arrows3D(x0=0,y0=0,z0=0,
         x1=vec[1], y1=vec[2], z1=vec[3], lwd = 2, col = color,
         add = TRUE)
}
```

### Apply the function
```{r}
vecplot_3d(vector = c(2, 3, 4), xax = c(0, 5), yax = c(0, 5), zax = c(0,5),
           col = "green")
```

2. Scalar multiplication

```{r}
v1 <- 1:5
v1
2 * v1
```
3. Vector addition
- Note that vectors must *all* be the same size
- R has built in functionality where if the longer vector is a multiple of the shorter vector length then it will just replicate the addition
  - Be careful not to let this happen accidentally!
```{r}
v2 <- c(-1, -2, -3, -4, -5)
v1 + v2
```

4. Vector subtraction
```{r}
v1 - v2
```
5. Scalar multiplication and vector addition together
```{r}
v1 + 5 * v2
```
6. Plotting vector addition

#### in 2-space
```{r}
vec.add.plot_2d <- function(vec1, vec2, xax, yax) {
  to.plot <- as.data.frame(rbind(vec1, vec2, vec1 + vec2))
  plot(to.plot[-2,]$v1, to.plot[-2,]$v2, type = 'p', col = "red",
       lwd = 2, xlim = xax, ylim = yax)
  
  # Add arrows
  arrows(x0 = 0, y0 = 0, x1 = to.plot[1,1], y1 = to.plot[1,2], lwd = 2, col = "blue")
  arrows(x0 = to.plot[1,1], y0 = to.plot[1,2], x1 = to.plot[3,1], y1 = to.plot[3,2], col = "green", lwd = 2)
  arrows(x0 = 0, y0 = 0, x1 = to.plot[3,1], y1 = to.plot[3,2], lwd = 2, col = "red")

}
```

### Apply the function
```{r}
vec.add.plot_2d(vec1 = c(1, 2), vec2 = c(3, 4), xax = c(0, 10), yax = c(0, 10))
```
  # Add point labels - removed - not working
  to.plot %>%
    mutate(point = paste('(', v1, ',', v2, ')', sep="")) -> to.plot
  with(to.plot[-2,], text(to.plot[-2,], labels = point, pos = 4))

### in 3-space
```{r}
library(plot3D)
vec.add.plot_3d <- function(vec1, vec2, xax, yax, zax, phi, theta) {
  points3D(x = vec1[1], y = vec1[2], z = vec1[3], col = 'red',
           xlim = xax, ylim = yax, zlim = zax,
           xlab = "x", ylab = "y", zlab = "z",
           phi = phi,
           theta = theta)
  text3D(x = vec1[1], y = vec1[2], z = vec1[3]+0.24,
         labels = paste("<",vec1[1],",",vec1[2],",",vec1[3],")", sep=""), add = TRUE)
  arrows3D(x0=0, y0=0, z0=0, x1 = vec1[1], y1 = vec1[2], z1 = vec1[3], lwd = 2, col = 'blue', add = TRUE)
  points3D(x = vec1[1] + vec2[1], y = vec1[2] + vec2[2], z = vec1[3]+vec2[3], col = 'green', xlim = xax, ylim = yax, zlim = zax,
           xlab = "x", ylab = "y", zlab = "z",
           add = TRUE)
  arrows3D(x0 = vec1[1], y0 = vec1[2], z0 = vec1[3],
           x1 = vec1[1] + vec2[1], y1 = vec1[2] + vec2[2], z1 = vec1[3] + vec2[3], lwd = 2, col = 'red', add = TRUE)
  
  vec3 <- vec1 + vec2
  points3D(x = vec3[1], y = vec3[2], z = vec3[4], col = 'red',
           xlim = xax, ylim = yax, zlim = zax,
           xlab = "x", ylab = "y", zlab = "z",
           add = TRUE)
  text3D(x = vec3[1], y = vec3[2], z = vec3[3],
         labels = paste("(", vec3[1],",",vec3[2],",",vec3[3],")",sep=""), add = TRUE)
  arrows3D(x0=0, y0=0, z0=0, x1=vec3[1], y1=vec3[2], z1=vec3[3], lwd = 2, col = "green", add = TRUE)

}
```

#### Apply function
```{r}
vec.add.plot_3d(vec1 = c(1,2,3), vec2 = c(4, 5, 6), xax = c(0, 10), yax = c(0, 10), zax = c(0,10), phi = 0, theta = 30)
```

## Section 5 - Lab2 - Intro to vector operations

1. Multiply vector v1, where v1 = [3, 5], by -3. What is the result?
-3*v1 = [-9, -15]
```{r}
v1 <- c(3, 5)
v1 * -3
```


2. Add v2 to v1 where v2 = [4, 2]
v1 + v2 = [7, 7]
```{r}
v2 <- c(4, 2)
v1 + v2
```

3. What is the sum of v2 and twice v1?
v2 + 2v1 = [4, 2] + 2*[3, 5]
= [10, 12]
```{r}
v2 + 2*v1
```

*By hand and programmatically*

4. Plot v1 and v2, along with their sum v1 + v2
```{r}
vec.add.plot_2d(vec1 = v1, vec2 = v2, 
                xax = c(0,10), yax = c(0, 10))
```


5. Add v3 = [0, 4, 3] and v4 = [4, 7, 5]
v3 + v4 = [4, 11, 8]
```{r}
v3 <- c(0, 4, 3)
v4 <- c(4, 7, 5)
v3 + v4
```

6. Plot v3 and v4 along with their sum v3 + v4. Change the perspective to get a better view of the vectors

```{r}
vec.add.plot_3d(vec1 = v3, vec2 = v4, 
                xax = c(0, 10), yax = c(0, 10), zax = c(0, 10),
                phi = 0, theta = 15)
```

## Section 6 - Matrix operations - Addition, subtraction and multiplication

#### Scalar multiplication of matrices
- Exactly like with vectors
  - Simply multiply each element of the matrix by the scalar

#### Matrix additiona and subtraction
- Exactly like with vectors
  - Dimensions must be exactly the same
  - Add or subtract element-wise
  - If begin with a matrix and subtract the same matrix - end with a 0 matrix (special case)
  
#### Matrix multiplication
- Only two matrices can be multiplied together at once
  - The *inner dimensions* of the two matrices *must* be exactly the same
    - E.g. a 2x4 matrix could be multiplied by a 4x3 matrix, but not by a 3x4
    - The number of cols in matrix 1, must equal the number of rows in the second matrix
    - Multiply element-wise using the rows from the first matrix and the columns from the second
      - Then add the row/column products
    - The dimensions of the product will be the *outer dimensions* of the two products
    - Multiply first row of matrix A by first col of matrix B and add together
      - Then the first row of matrix A by second col of matrix B and sum
      - And repeat until reach last col

- **Vector multiplication**
  - Must be a row vector by a column vector, or column times row
  - 1xn times nx1 or nx1 times 1xn
  - Returns a scalar if 1xn times nx1
  
- Matrix multiplication is not commutative
  - **order matters!**
  - A * B != B * A
  
## Section 7 - Application - matrix operations, addition, subtraction, multiplication

1. Scalar multiplication of matrices
```{r}
A <- matrix(c(-1, 2, 7, 0), nrow = 2, ncol = 2, byrow = TRUE)
A
2*A
```

2. Matrix addition and subtraction
- Non-conformable arrays - dimensionality isn't correct
```{r}
A <- matrix(c(0, 1, 2, 3), nrow = 2, ncol = 2, byrow = TRUE)
B <- matrix(c(0, -2, -5, 12), nrow = 2, ncol = 2, byrow = TRUE)
A
B
A+B

A2 <- matrix(1:6, nrow = 2, byrow = TRUE)
A + A2

C <- matrix(c(-3, 0, 1, 7, 4, -6), nrow = 3, byrow = TRUE)
D <- matrix(c(-3, 0, 1, 7, 4, -6), nrow = 3, byrow = TRUE)

C - D
```


3. Matrix multiplication
- Need to be careful with R commands
  - A * B -> doesn't give the right answer
    - **Element-wise multiplication** - not true multiplication
  - A %*% B is what you want
```{r}
A
B
A * B
A %*% B
# Both 2x2, expect a 2x2 answer

C # 3x2
D # 3x2
# Can't do C %*% D

B %*% A
```

## Section 8 - LAB3 - Intro matrix operations addition subtraction and multiplication

1. Define two matrices. Let A by 3x4, filled row-wise with numbers 1-12. Let B be 4x2, filled column-wise with the numbers 1-8
```{r}
A <- matrix(1:12, nrow = 3, byrow = TRUE)
B <- matrix(1:8, nrow = 4, byrow = TRUE)
A
B
```


2. Explain why A and B cannot be added or subtracted.
A and B cannot be added or subtracted because their dimensions are not the same

3. Create matrix C that can be added to A. Perform the addition.
```{r}
C <- matrix(13:24, nrow = 3, byrow = TRUE)
C
A + C
```


4. Find AxB before multiplying in R, anticipate the dimensions of the product, and multiply by hand
```{r}
A %*% B
```

