---
title: "Linear Algebra"
output: html_notebook
---

## Section 1 - Scalars Vectors and Matrices

- Scalar - a constant
- Vector - a single row or column of numbers
- Matrix - a rectangular collection of numbers

- Dimensions - refer to the number of rows and columns
  - an rxc matrix - r rows and c columns
  - The m~ij~ element is at row i and column j
  
- Vectors are a subset of matrices
  - Either dimensions 1xc (a row vector) or rx1 (a column vector)
  - And scalars are 1x1
  
- Vectors are used to represent a location in space. Each entry represents a dimension. This is easy to see in 1, 2 and 3 dimensions
- Vectors are arrows, whenever we get beyond a single dimension

## Section 2 - Application Scalars Vectors and Metrics

1. Defining a vector [c(), :, seq(), rep()]
- R is fully vectorized - default process is to think in terms of vectors (i.e. a collection of elements)

```{r}
# Using the c() function
v1 <- c(2, 3, 4, 5)
str(v1)
# Using the ':' operator
v2 <- 2:5 # From 2 to 5, default counter 1

# USing the seq() function; a long-form version of the ':' operator
v3 <- seq(from = 2, to = 5, by = 1)

# OR...
v3 <- seq(2, 5, 1)

# Using the rep() function
v4 <- rep(37, 5)  # Repeats 37 5x times
```

2. Defining a matrix [matrix(), rbind(), cbind()]
- A matrix is just a collection of vectors
```{r}
# Using the matrix() function
m1 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = TRUE)
# byrow = TRUE - takes arguments and populates rowwise
m1
# Default is FALSE
# Useful to do it rowwise so can think rows first
str(m1)

# Using rbind()
m2 <- rbind(c(1, 2, 3), c(4, 5, 6))
m2
# Using cbind()
m3 <- cbind(c(1, 4), c(2, 5), c(3, 6))
m3
```

3. Plots in 2D and 3D
- Vector in 2-space
```{r}
# Vector in 2-space
u1 <- c(2, 4)
plot(u1[1], u1[2], type = 'p', col = 'red',
     xlim = c(0, 5), ylim = c(0,5),
     xlab = "", ylab = "")  
arrows(x0 = 0, y0 = 0,
       x1 = u1[1], y1 = u1[2],
       lwd = 2)  
# *0 - Tells where arrows should start
# Use reference to tell plot elements
```

- Vector in 3-space
```{r}
library(plot3D)
u2 <- c(2, 3, 4)
points3D(x=u2[1], y=u2[2], z=u2[3], col = 'red',
         xlim = c(0,5), ylim = c(0,5), zlim = c(0,5),
         phi = 0)  # Changes view
arrows3D(x0 = 0, y0 = 0, z0 = 0, x1 = u2[1], y1 = u2[2], z1 = u2[3],
         lwd = 2,
         add = TRUE)  # Adds to an existing plot - needed for 3D library
```

## Section 2 - LAB1 - Intro Scalars Vectors and Matrices
1. Create a 3x4 matrix, A, fill row-wise with the number 1-12. What is a~24~?
```{r}
A <- matrix(seq(1,12,1), nrow = 3, ncol = 4, byrow = TRUE)
A
A[2,4]
# a_24 = 8
```

2. Create two 1x2 vectors, v1 and v2, with numeric entities of your choice. Plot v1 and v2 on the same coordinate plane
```{r}
v1 <- c(5, 6)
v2 <- c(1, 3)
# Plot begins a session
plot(v1[1], v1[2], type = 'p', col = 'red',
     xlim = c(0,8), ylim = c(0,8),
     xlab = "", ylab = "")
arrows(x0 = 0, y0 = 0,
       x1 = v1[1], y1 = v1[2],
       lwd = 2, col = 'red')
# Points adds additional points
points(v2[1], v2[2], col = 'blue',
     xlim = c(0,8), ylim = c(0,8),
     xlab = "", ylab = "")

arrows(x0 = 0, y0 = 0,
       x1 = v2[1], y1 = v2[2],
       lwd = 2, col = 'blue')
```

3. Now add 3 to the second entry in v1 and re-plot. Describe the position of the new vector in 2-space relative to v1
```{r}
v2_b <- c(v1[1], v1[2] +3)
v2_b
plot(v1[1], v1[2], type = 'p', col = 'red',
     xlim = c(0,10), ylim = c(0,10),
     xlab = "", ylab = "")
plot(v2_b[1], v2_b[2], type = 'p', col = 'blue',
     xlim = c(0,10), ylim = c(0,10),
     xlab = "", ylab = "")
arrows(x0 = 0, y0 = 0,
       x1 = v1[1], y1 = v1[2],
       lwd = 2, col = 'red')
arrows(x0 = 0, y0 = 0,
       x1 = v2_b[1], y1 = v2_b[2],
       lwd = 2, col = 'blue')
```

4. Create a 1x3 vector, v3, with numeric entries of your choice. Plot v3 in 3-space
```{r}
v3 <- c(4, 5, 6)
points3D(x=v3[1], y = v3[2], z = v3[3], col = "red",
         xlim = c(0,8), ylim = c(0,8), zlim = c(0,8))
arrows3D(x0 = 0, y0 = 0, z0 = 0,
         x1 = v3[1], y1 = v3[2], z1 = v3[3],
         lwd = 2, add = TRUE)
```

5. Fill in the blanks: A *vector* is a *matrix* with only one *row* or *column*

## Section 3 - Vector operations
- **Scalar multiplication** - a scalar times a vector
  - Multiply each element of the vector by the scalar
  - 3 x [-1, 2, 5] = [-3. 6, 15]
  - The original vector is being **dilated** (stretched/compressed) by a factor equal to the scalar
  - If scalar > 1, then stretched
  - If scalar < 1, then compressed
  - If <0, then the direction will be flipped
  
- **Vector addition** - adding two vectors
  - Both vectors must have exactly the same dimensions
  - Then we add element wise
  - e.g. [2, 3] + [1, 4] = [3, 7]
  - The second vector is being drawn from the end of the first vector to create a new vector
  
- **Vector subtraction** - simply add the negative vector
  - Multiply each element of the second vector by the scalar -1, then subtract
  - v1 - v2 = v1 + -1*v2
  - Net effect is the same as subtracting the second vector from the first element-wise
  - r.h. [2, 3] - [1, 4] = [1, -1]
  - Illustrates a more general rule:
    - Multiplication and addition can be applied successively
    - v1 + s*v2
    - The order of operations is the same as with real number - multiply first, then add/subtract
    - Multiple each element of the second vector by the scalar then subtract
    - e.g. [2, 3] + 3 * [1, 4] = [5, 15]
    
- Linear algebra - in general, don't assume that rules in the real-number world still apply. The above is an exception

## Section 4 - Application vector operations

1. Plotting vectors
#### Plotting a single vector in 2-space

```{r}
vecplot_2d <- function(vector, xax, yax, color) {
  vec <- vector
  plot(vec[1], vec[2], type = 'p', col = color,
  xlim = xax, ylim = yax,
  xlab = "", ylab = "")
  arrows(x0 = 0, y0 = 0, x1 = vec[1], y1 = vec[2], lwd = 2, col = color)
  text(x = vec[1], y = vec[2], 
  labels = paste('(', vec[1], ',', vec[2], ')', sep=""),
  pos = 3)  # Pos 3 is to the right
  }
```

#### Apply the function
```{r}
vecplot_2d(vector = c(1,2), xax = c(0,4), yax = c(0,5), color="red")
```

#### Plotting a vector in 3-space
```{r}
library(plot3D)
vecplot_3d <- function(vector, xax, yax, zax, color) {
  vec <- vector
  points3D(x = vec[1], y = vec[2], z = vec[3],
           xlim = xax, ylim = yax, zlim = zax,
    xlab = "x", ylab = "y", zlab = "z")
text3D(x = vec[1] + 0.5, y = vec[2], z = vec[3],
       labels = paste("(",vec[1], ",", vec[2], ",", vec[3],")", sep = ""),
       add = TRUE)
arrows3D(x0=0,y0=0,z0=0,
         x1=vec[1], y1=vec[2], z1=vec[3], lwd = 2, col = color,
         add = TRUE)
}
```

### Apply the function
```{r}
vecplot_3d(vector = c(2, 3, 4), xax = c(0, 5), yax = c(0, 5), zax = c(0,5),
           col = "green")
```

2. Scalar multiplication

```{r}
v1 <- 1:5
v1
2 * v1
```
3. Vector addition
- Note that vectors must *all* be the same size
- R has built in functionality where if the longer vector is a multiple of the shorter vector length then it will just replicate the addition
  - Be careful not to let this happen accidentally!
```{r}
v2 <- c(-1, -2, -3, -4, -5)
v1 + v2
```

4. Vector subtraction
```{r}
v1 - v2
```
5. Scalar multiplication and vector addition together
```{r}
v1 + 5 * v2
```
6. Plotting vector addition

#### in 2-space
```{r}
vec.add.plot_2d <- function(vec1, vec2, xax, yax) {
  to.plot <- as.data.frame(rbind(vec1, vec2, vec1 + vec2))
  plot(to.plot[-2,]$v1, to.plot[-2,]$v2, type = 'p', col = "red",
       lwd = 2, xlim = xax, ylim = yax)
  
  # Add arrows
  arrows(x0 = 0, y0 = 0, x1 = to.plot[1,1], y1 = to.plot[1,2], lwd = 2, col = "blue")
  arrows(x0 = to.plot[1,1], y0 = to.plot[1,2], x1 = to.plot[3,1], y1 = to.plot[3,2], col = "green", lwd = 2)
  arrows(x0 = 0, y0 = 0, x1 = to.plot[3,1], y1 = to.plot[3,2], lwd = 2, col = "red")

}
```

### Apply the function
```{r}
vec.add.plot_2d(vec1 = c(1, 2), vec2 = c(3, 4), xax = c(0, 10), yax = c(0, 10))
```
  # Add point labels - removed - not working
  to.plot %>%
    mutate(point = paste('(', v1, ',', v2, ')', sep="")) -> to.plot
  with(to.plot[-2,], text(to.plot[-2,], labels = point, pos = 4))

### in 3-space
```{r}
library(plot3D)
vec.add.plot_3d <- function(vec1, vec2, xax, yax, zax, phi, theta) {
  points3D(x = vec1[1], y = vec1[2], z = vec1[3], col = 'red',
           xlim = xax, ylim = yax, zlim = zax,
           xlab = "x", ylab = "y", zlab = "z",
           phi = phi,
           theta = theta)
  text3D(x = vec1[1], y = vec1[2], z = vec1[3]+0.24,
         labels = paste("<",vec1[1],",",vec1[2],",",vec1[3],")", sep=""), add = TRUE)
  arrows3D(x0=0, y0=0, z0=0, x1 = vec1[1], y1 = vec1[2], z1 = vec1[3], lwd = 2, col = 'blue', add = TRUE)
  points3D(x = vec1[1] + vec2[1], y = vec1[2] + vec2[2], z = vec1[3]+vec2[3], col = 'green', xlim = xax, ylim = yax, zlim = zax,
           xlab = "x", ylab = "y", zlab = "z",
           add = TRUE)
  arrows3D(x0 = vec1[1], y0 = vec1[2], z0 = vec1[3],
           x1 = vec1[1] + vec2[1], y1 = vec1[2] + vec2[2], z1 = vec1[3] + vec2[3], lwd = 2, col = 'red', add = TRUE)
  
  vec3 <- vec1 + vec2
  points3D(x = vec3[1], y = vec3[2], z = vec3[4], col = 'red',
           xlim = xax, ylim = yax, zlim = zax,
           xlab = "x", ylab = "y", zlab = "z",
           add = TRUE)
  text3D(x = vec3[1], y = vec3[2], z = vec3[3],
         labels = paste("(", vec3[1],",",vec3[2],",",vec3[3],")",sep=""), add = TRUE)
  arrows3D(x0=0, y0=0, z0=0, x1=vec3[1], y1=vec3[2], z1=vec3[3], lwd = 2, col = "green", add = TRUE)

}
```

#### Apply function
```{r}
vec.add.plot_3d(vec1 = c(1,2,3), vec2 = c(4, 5, 6), xax = c(0, 10), yax = c(0, 10), zax = c(0,10), phi = 0, theta = 30)
```

## Section 5 - Lab2 - Intro to vector operations

1. Multiply vector v1, where v1 = [3, 5], by -3. What is the result?
-3*v1 = [-9, -15]
```{r}
v1 <- c(3, 5)
v1 * -3
```


2. Add v2 to v1 where v2 = [4, 2]
v1 + v2 = [7, 7]
```{r}
v2 <- c(4, 2)
v1 + v2
```

3. What is the sum of v2 and twice v1?
v2 + 2v1 = [4, 2] + 2*[3, 5]
= [10, 12]
```{r}
v2 + 2*v1
```

*By hand and programmatically*

4. Plot v1 and v2, along with their sum v1 + v2
```{r}
vec.add.plot_2d(vec1 = v1, vec2 = v2, 
                xax = c(0,10), yax = c(0, 10))
```


5. Add v3 = [0, 4, 3] and v4 = [4, 7, 5]
v3 + v4 = [4, 11, 8]
```{r}
v3 <- c(0, 4, 3)
v4 <- c(4, 7, 5)
v3 + v4
```

6. Plot v3 and v4 along with their sum v3 + v4. Change the perspective to get a better view of the vectors

```{r}
vec.add.plot_3d(vec1 = v3, vec2 = v4, 
                xax = c(0, 10), yax = c(0, 10), zax = c(0, 10),
                phi = 0, theta = 15)
```

## Section 6 - Matrix operations - Addition, subtraction and multiplication

#### Scalar multiplication of matrices
- Exactly like with vectors
  - Simply multiply each element of the matrix by the scalar

#### Matrix additiona and subtraction
- Exactly like with vectors
  - Dimensions must be exactly the same
  - Add or subtract element-wise
  - If begin with a matrix and subtract the same matrix - end with a 0 matrix (special case)
  
#### Matrix multiplication
- Only two matrices can be multiplied together at once
  - The *inner dimensions* of the two matrices *must* be exactly the same
    - E.g. a 2x4 matrix could be multiplied by a 4x3 matrix, but not by a 3x4
    - The number of cols in matrix 1, must equal the number of rows in the second matrix
    - Multiply element-wise using the rows from the first matrix and the columns from the second
      - Then add the row/column products
    - The dimensions of the product will be the *outer dimensions* of the two products
    - Multiply first row of matrix A by first col of matrix B and add together
      - Then the first row of matrix A by second col of matrix B and sum
      - And repeat until reach last col

- **Vector multiplication**
  - Must be a row vector by a column vector, or column times row
  - 1xn times nx1 or nx1 times 1xn
  - Returns a scalar if 1xn times nx1
  
- Matrix multiplication is not commutative
  - **order matters!**
  - A * B != B * A
  
## Section 7 - Application - matrix operations, addition, subtraction, multiplication

1. Scalar multiplication of matrices
```{r}
A <- matrix(c(-1, 2, 7, 0), nrow = 2, ncol = 2, byrow = TRUE)
A
2*A
```

2. Matrix addition and subtraction
- Non-conformable arrays - dimensionality isn't correct
```{r}
A <- matrix(c(0, 1, 2, 3), nrow = 2, ncol = 2, byrow = TRUE)
B <- matrix(c(0, -2, -5, 12), nrow = 2, ncol = 2, byrow = TRUE)
A
B
A+B

A2 <- matrix(1:6, nrow = 2, byrow = TRUE)
A + A2

C <- matrix(c(-3, 0, 1, 7, 4, -6), nrow = 3, byrow = TRUE)
D <- matrix(c(-3, 0, 1, 7, 4, -6), nrow = 3, byrow = TRUE)

C - D
```


3. Matrix multiplication
- Need to be careful with R commands
  - A * B -> doesn't give the right answer
    - **Element-wise multiplication** - not true multiplication
  - A %*% B is what you want
```{r}
A
B
A * B
A %*% B
# Both 2x2, expect a 2x2 answer

C # 3x2
D # 3x2
# Can't do C %*% D

B %*% A
```

## Section 8 - LAB3 - Intro matrix operations addition subtraction and multiplication

1. Define two matrices. Let A by 3x4, filled row-wise with numbers 1-12. Let B be 4x2, filled column-wise with the numbers 1-8
```{r}
A <- matrix(1:12, nrow = 3, byrow = TRUE)
B <- matrix(1:8, nrow = 4, byrow = TRUE)
A
B
```


2. Explain why A and B cannot be added or subtracted.
A and B cannot be added or subtracted because their dimensions are not the same

3. Create matrix C that can be added to A. Perform the addition.
```{r}
C <- matrix(13:24, nrow = 3, byrow = TRUE)
C
A + C
```


4. Find AxB before multiplying in R, anticipate the dimensions of the product, and multiply by hand
```{r}
A %*% B
```

## Section 8 - Matrix operations - matrix transposes and inverses
- **Transpose** - all of the elements switch positions
  - Switch row/col position between every element
  - If A is a 3X5 then A^T^ is a 5x3 matrix
- **Inverse** of a matrix
  - The matrix *must* be square
  - Cannot have a row or column of all zeros
  - Another condition exists - but it is more technical. Won't go into here, but quite rare
  - If all satisfied - matrix is *invertible*
  
- For 2x2 matrix A (matrix(c(a, b, c, d), nrows = 2, byrow = TRUE))
- All four elements are quotients
  - Denominator is the product of one diag minus product of other diag
  - Numerator
    - a and d have swapped places
    - b and c stay in the same place, but have been negated
  - As dimensionality increases - the process becomes more complex

- When a matrix is multiplied by its invese A*A^-1^
  - Returns an **identity matrix**
    - Any nxn matrix whose main diagonal is all 1s and has 0s everywhere else
    - Any matrix times its corresponding identity matrix will yield itself
      A*I = A (works like 1 in real number algebra)
    - Any matrix times its inverse yields the corresponding identity matrix
      A*A^-1^ = I (analagous to any number times its reciprocal in real number algebra)

## Section 9 - Application - Matrix transposes and inverses

1. Matrix transpose
- The t function is handy, but needs a matrix type object passed to it
- It doesn't work that well with a vector
- Always pass t() a matrix object
```{r}
A <- matrix(1:12, nrow = 3, ncol = 2, byrow = TRUE)
A

A.t <- t(A)  # t() transposes matrix
A.t

v1 <- matrix(1:4, nrow=1)
v1

v1.t <- t(v1)
v1.t
```

2. Matrix inverse
- `A^-1` does not return the inverse - returns element wise reciprocation
- I.e. taking each element and dividing by 1

- If we want the inverse - `solve(A)`

- Don't forget - only square matrices are invertible
- Matrix can not have a row or column of 0's
  - Returns - the system is exactly singular
  - Means that the denominator is 0
```{r}
A <- matrix(1:4, nrow=2, byrow = TRUE)
A

A^-1 # Not what we expected!
solve(A) # Produces the inverse

B <- matrix(1:6, nrow = 2)
solve(B)

C <- matrix(c(1,2,0,0), nrow = 2)
C
solve(C)
```

3. Identity matrix
- A matrix times its inverse gives the identity
```{r}
A %*% solve(A)
# Machine error in both entries in col 2 - results in scientific error
# Effectively an identity matrix
# If we were to round then it comes out ok
round(A %*% solve(A))
```

- Get the identity directly using `diag()`
```{r}
I2 <- diag(2)
I2

I3 <- diag(3)
I3

I7 <- diag(7)
I7
```

- A matrix times the identity yields the matrix again
```{r}
A %*% I2
```

## Section 10 - Lab 4 - Intro Matrix Operations Transposes and Inverses

1. Define matrix A as a square matrix with entries of your choice.
```{r}
A <- matrix(1:9, nrow = 3, byrow = TRUE)
A
```

2. Find the transpose of A - by hand and in R
```{r}
A.t <- t(A)
A.t
```

3. Find the inverse of A, if it exists. If it does not exist, define a new matrix A that is invertible
```{r}
A.inv = solve(A)
A.inv

A4 <- matrix(c(5, 2, 1, 7, 4, 8, 6, 3, 2), nrow = 3, byrow = TRUE)
A4

A4.inv = solve(A4)
A4.inv
```


4. Multiply A by its inverse. Before - what is the expected result?
```{r}
round(A4 %*% A4.inv)
```

## Section 11 - What is linear regression
1. Definition and Generic Model Statement
- **Linear regression** - method of creating a *linear equation* to model the relationship between a dependent variable and one or more independent variable

2. Deterministic example
- Suppose you're going to a baseball game.
  - Ticket cost = $20
  - Hot dogs $3
  - French fries $2
- Your total cost can be modelled as a linear function of the ticket price and the number of hot dogs and French fries you purchase
  - Total cost equals $20, plus $3 for every hot dog and $2 for every order of fries
  - Mathematically: c = 20 + 3h + 2f
  - h = no hot dogs, f = number of fries
  - Linear regression comes in when we only have the data and need to estimate the coefficients in the equation

3. Stochastic example
```{r}
mtcars
```
- Suppose you wish to model *mpg* as a function of *wt* and *qsec*
  - The coefficients are unknown
  - Via linear regression the model is:
  mpg = 19.75 + (-5.05)*wt + 0.93 * qsec + error
  - Coeffs - how much of a contribution to the independent variable do the dependent variables make

## Section 12 - Application - what is linear regression

1. Deterministic example
- Var1 -> hotdogs
- Var2 -> fries
- game-cost -> shows all different possible combinations of fries and hot.dogs
```{r}
library(tidyverse)

# Create hot dog data and plot
hot.dogs <- 0:20  # The number of hot dogs purchased
fries <- 0:20  # The number of fries purchased
game.cost <- expand.grid(hot.dogs, fries)
# Creates a cross-product of the two variables
game.cost
colnames(game.cost) <- c('Hot.dogs', 'Fries')
game.cost <- game.cost %>%
  mutate(Ticket = 20,
         Total.Cost = Ticket + 3*Hot.dogs + 2*Fries)


### Create a plot
library(rgl)
plot3d(x = game.cost$Hot.dogs,
       y = game.cost$Fries,
       z = game.cost$Total.Cost,
       type = 'p',
       col = 'green',
       xlim = c(0,20),
       ylim = c(0,20),
       zlim = c(0,120),
       xlab = "Hot Dogs purchased",
       ylab = "French Fries purchased",
       zlab = "Total Cost")

### Use lm() function (linear model) to estimate a linear model
hot.dog.model <- lm(Total.Cost ~ Hot.dogs + Fries, data=game.cost)
hot.dog.model
summary(hot.dog.model)

```
- Estimate column - shows the coefficients

2. Stochastic example
```{r}
view(mtcars)  # Opens up df in separate tab

### Create a plot
plot3d(x = mtcars$wt,
       y = mtcars$qsec,
       z = mtcars$mpg,
       type = 's',
       col = 'green',
       xlim = c(min(mtcars$wt)-1, max(mtcars$wt)+1),
       ylim = c(min(mtcars$qsec)-1, max(mtcars$qsec)+1),
       zlim = c(min(mtcars$mpg)-1, max(mtcars$mpg)+1),
       xlab = 'Wt',
       ylab = 'Qsec',
       zlab = 'MPG')

# Use lm() to fit a linear regression  model
mpg.model <- lm(mpg ~ wt + qsec, data = mtcars)
mpg.model

# Plot model vs. actuals
plot(mtcars$mpg, mpg.model$fitted.values)  # Want a line
# Fitted values calculated from the regression formula

# Plot model vs. actuals
plot3d(x = mtcars$wt,
       y = mtcars$qsec,
       z = mpg.model$fitted.values,
       type = "s",
       col = "red",
       xlim = c(min(mtcars$wt)-1, max(mtcars$wt)+1),
       ylim = c(min(mtcars$qsec)-1, max(mtcars$qsec)+1),
       zlim = c(min(mtcars$mpg)-1, max(mtcars$mpg)+1),
       xlab = "Wt",
       ylab = "Qsec",
       zlab = "MPG",
       add = TRUE)

```

3. Model error
```{r}
# Different outputs
mpg.model # Gives you the 'beta' vector
summary(mpg.model) # Gives you more detailed output
```
- Calculating error measures for the model
```{r}
error.measures <- as.data.frame(cbind(mtcars$mpg, mpg.model$fitted.values))
colnames(error.measures) <- c('Actuals', 'Predicted.Values')
error.measures <- error.measures %>%
  mutate(error = Actuals - Predicted.Values,
         squared.error = error ^ 2)
error.measures

# Why square the errors?
sum(error.measures$error)  # Essentially 0
# Regression equation fits perfectly balanced through the equation data
# To get around this - we square the errors
boxplot(error.measures$error)  # Evenly distributed
boxplot(error.measures$squared.error)  # Knocked things out of scale
# Advantage - get rid of negatives
# Disadvantage - knock everything out of scale

# mse and rmse
mean.squared.error <- mean(error.measures$squared.error)
mean.squared.error

root.mse <- sqrt(mean.squared.error)
root.mse
# root.mse brings squared error back into original scale
# RMSE is the measure of total error in the model
# On average - predicted value would be about 2.5 units out in one direction or other
```
# Section 13 - Lab 5 - Intro to What is Linear Regression

1. Download the 'Baseball2018.csv' dataset and load it into R using read.csv()
- No access to 'Baseball2018.csv'
```{r}
mtcars

```


2. Plot *Runs* vs. *HomeRun* and *Runs* vs. *H9*. Of *HomeRun* and *H9*, which appears to be more closely related to *Runs*? Plot *Runs* vs. *HomeRun* and *H9* in 3-space
```{r}
# Independent variable - mpg
# Dependent variables - hp (horsepower), disp (displacement)

# Plot mpg and horsepower
plot(mtcars$hp, mtcars$mpg) # Negative relationship (non linear?)
# Plot mpg and displacement
plot(mtcars$disp, mtcars$mpg) # Also negative relationship
# 3D plot mpg, disp and hp
```


3. Use the `lm()` function to estimate the linear regression model for *Runs* as a function of *HomeRun* and *H9*. Write the regression equation. Plot fitted values against actuals. How accurate would you say the predictions are?
4. Calculate the RMSE for the model and offer an interpretation for the value in the context of the problem. Does this change your perception of the quality of the model predictions?

