---
title: "Linear Algebra"
output: html_notebook
---

## Section 1 - Scalars Vectors and Matrices

- Scalar - a constant
- Vector - a single row or column of numbers
- Matrix - a rectangular collection of numbers

- Dimensions - refer to the number of rows and columns
  - an rxc matrix - r rows and c columns
  - The m~ij~ element is at row i and column j
  
- Vectors are a subset of matrices
  - Either dimensions 1xc (a row vector) or rx1 (a column vector)
  - And scalars are 1x1
  
- Vectors are used to represent a location in space. Each entry represents a dimension. This is easy to see in 1, 2 and 3 dimensions
- Vectors are arrows, whenever we get beyond a single dimension

## Section 2 - Application Scalars Vectors and Metrics

1. Defining a vector [c(), :, seq(), rep()]
- R is fully vectorized - default process is to think in terms of vectors (i.e. a collection of elements)

```{r}
# Using the c() function
v1 <- c(2, 3, 4, 5)
str(v1)
# Using the ':' operator
v2 <- 2:5 # From 2 to 5, default counter 1

# USing the seq() function; a long-form version of the ':' operator
v3 <- seq(from = 2, to = 5, by = 1)

# OR...
v3 <- seq(2, 5, 1)

# Using the rep() function
v4 <- rep(37, 5)  # Repeats 37 5x times
```

2. Defining a matrix [matrix(), rbind(), cbind()]
- A matrix is just a collection of vectors
```{r}
# Using the matrix() function
m1 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = TRUE)
# byrow = TRUE - takes arguments and populates rowwise
m1
# Default is FALSE
# Useful to do it rowwise so can think rows first
str(m1)

# Using rbind()
m2 <- rbind(c(1, 2, 3), c(4, 5, 6))
m2
# Using cbind()
m3 <- cbind(c(1, 4), c(2, 5), c(3, 6))
m3
```

3. Plots in 2D and 3D
- Vector in 2-space
```{r}
# Vector in 2-space
u1 <- c(2, 4)
plot(u1[1], u1[2], type = 'p', col = 'red',
     xlim = c(0, 5), ylim = c(0,5),
     xlab = "", ylab = "")  
arrows(x0 = 0, y0 = 0,
       x1 = u1[1], y1 = u1[2],
       lwd = 2)  
# *0 - Tells where arrows should start
# Use reference to tell plot elements
```

- Vector in 3-space
```{r}
library(plot3D)
u2 <- c(2, 3, 4)
points3D(x=u2[1], y=u2[2], z=u2[3], col = 'red',
         xlim = c(0,5), ylim = c(0,5), zlim = c(0,5),
         phi = 0)  # Changes view
arrows3D(x0 = 0, y0 = 0, z0 = 0, x1 = u2[1], y1 = u2[2], z1 = u2[3],
         lwd = 2,
         add = TRUE)  # Adds to an existing plot - needed for 3D library
```

## Section 2 - LAB1 - Intro Scalars Vectors and Matrices
1. Create a 3x4 matrix, A, fill row-wise with the number 1-12. What is a~24~?
```{r}
A <- matrix(seq(1,12,1), nrow = 3, ncol = 4, byrow = TRUE)
A
A[2,4]
# a_24 = 8
```

2. Create two 1x2 vectors, v1 and v2, with numeric entities of your choice. Plot v1 and v2 on the same coordinate plane
```{r}
v1 <- c(5, 6)
v2 <- c(1, 3)
# Plot begins a session
plot(v1[1], v1[2], type = 'p', col = 'red',
     xlim = c(0,8), ylim = c(0,8),
     xlab = "", ylab = "")
arrows(x0 = 0, y0 = 0,
       x1 = v1[1], y1 = v1[2],
       lwd = 2, col = 'red')
# Points adds additional points
points(v2[1], v2[2], col = 'blue',
     xlim = c(0,8), ylim = c(0,8),
     xlab = "", ylab = "")

arrows(x0 = 0, y0 = 0,
       x1 = v2[1], y1 = v2[2],
       lwd = 2, col = 'blue')
```

3. Now add 3 to the second entry in v1 and re-plot. Describe the position of the new vector in 2-space relative to v1
```{r}
v2_b <- c(v1[1], v1[2] +3)
v2_b
plot(v1[1], v1[2], type = 'p', col = 'red',
     xlim = c(0,10), ylim = c(0,10),
     xlab = "", ylab = "")
plot(v2_b[1], v2_b[2], type = 'p', col = 'blue',
     xlim = c(0,10), ylim = c(0,10),
     xlab = "", ylab = "")
arrows(x0 = 0, y0 = 0,
       x1 = v1[1], y1 = v1[2],
       lwd = 2, col = 'red')
arrows(x0 = 0, y0 = 0,
       x1 = v2_b[1], y1 = v2_b[2],
       lwd = 2, col = 'blue')
```

4. Create a 1x3 vector, v3, with numeric entries of your choice. Plot v3 in 3-space
```{r}
v3 <- c(4, 5, 6)
points3D(x=v3[1], y = v3[2], z = v3[3], col = "red",
         xlim = c(0,8), ylim = c(0,8), zlim = c(0,8))
arrows3D(x0 = 0, y0 = 0, z0 = 0,
         x1 = v3[1], y1 = v3[2], z1 = v3[3],
         lwd = 2, add = TRUE)
```

5. Fill in the blanks: A *vector* is a *matrix* with only one *row* or *column*

## Section 3 - Vector operations
- **Scalar multiplication** - a scalar times a vector
  - Multiply each element of the vector by the scalar
  - 3 x [-1, 2, 5] = [-3. 6, 15]
  - The original vector is being **dilated** (stretched/compressed) by a factor equal to the scalar
  - If scalar > 1, then stretched
  - If scalar < 1, then compressed
  - If <0, then the direction will be flipped
  
- **Vector addition** - adding two vectors
  - Both vectors must have exactly the same dimensions
  - Then we add element wise
  - e.g. [2, 3] + [1, 4] = [3, 7]
  - The second vector is being drawn from the end of the first vector to create a new vector

