---
title: "Overview of R"
output: html_notebook
---

#### Overview of R workspace and basic commands


```{r}
data.00 <- mtcars
```

#### Tidyverse standards and data wrangling

```{r}
library(tidyverse)
```


1. Every column is a unique variable
2. Every row is a unique observation
3. Only keep what you need in each dataset

# Working with what you have
- Filter for where cyl is 4
```{r}
# Filter for where cyl is 4
filter(data.00, cyl == 4)
# Subset columns
select(data.00, mpg, cyl)

data.00 %>% # Piping operator
  filter(cyl == 4)

data.00 %>%
  select(mpg, cyl)

### Creating something new
data.00 %>%
  mutate(mpg.per.cyl = mpg / cyl) %>%
  group_by(cyl) %>%
  summarize(mean.mpg.per.cyl = mean(mpg.per.cyl))

# Quick plot
plot(x = data.00$cyl, y = data.00$mpg, xlab = "Cylinders", ylab = "Miles per Gallon", col = "red", main = "MPG by number of Cylinders")
```


### Lab
# Load the mpg dataset
```{r}
mpg
```
1. View the first 6 observations in two different ways
```{r}
head(mpg, 6)
mpg[1:6,]
slice(data.00, 1:6)
str(mpg)
```

2. Find the average of 'cty' for compact cars?
```{r}
mpg %>%
  group_by(class) %>%
  summarize(mean_cty = mean(cty)) %>%
  filter(class == 'compact')

# Alternatively
mpg %>%
  select(class, cty) %>%  # Subset only desired variables
  group_by(class) %>% # Collect all vehicles of the same class in the calculation
  summarize(mean.cty = mean(cty)) %>%
  filter(class == 'compact')
```


3. Create a new df with 5 variables: manufacturer, class, cyl, mean.hyw and hwy.per.cyl
  - mean.hwy should be the average of 'hwy' for each manufacturer's classes of vehicles, by number of cylinders - group by manuf, class and no of cyl
  - hwy.per.cyl should be the average ratio of hwy to cyl for each manufacturer's classes of vehicles, by number of cylinders
  - Each manufacturer/class/cyl combination should have only one record in the dataframe
  
```{r}
mpg %>%
  mutate(hwy.per.cyl = hwy / cyl) %>%
  group_by(manufacturer, class, cyl) %>% # Creates working groups
  summarize(mean.hwy = mean(hwy),
            hwy.per.cyl = mean(hwy.per.cyl)) %>%
  select(manufacturer, class, cyl, mean.hwy, hwy.per.cyl)
```

  
4. Plot 'hwy' against 'cty' for all 2008 models. Make the points blue 
```{r}
models_2008 <- mpg %>%
  filter(year == 2008)

plot(x = models_2008$hwy, y = models_2008$cty, type = 'p', col = "blue", main = "Hwy vs. cty for 2008 models", xlab = "Highway", ylab = "City")
```

  